---
title: "Class 11: Structural Bioinformatics"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##The PDB Database 

The [PDB](http://www.rcsb.org/) database is main repository for biomolecular structure data 

Here we examine the contents of the PDB:

>Q1: Download a CSV file from the PDB site (accessible from “Analyze” -> “PDB Statistics” >
“by Experimental Method and Molecular Type”. Move this CSV file into your RStudio project
and determine the percentage of structures solved by X-Ray and Electron Microscopy. From
the website what proportion of structures are protein?

```{r}
db <- read.csv("Data Export Summary.csv", row.names = 1)
head(db)

```

How many are Xray, etc... ( by percentage)

```{r}
(db$Total/sum(db$Total)) *100
```

What percentare protein...
```{r}
(sum(db$Proteins)/sum(db$Total) )*100
```

We can also try the datapasta package and copy from website (not a csv file, but just normal table) and  "Addins" > "Paste as data.frame"
```{r}
library(datapasta)

tmp <- data.frame(stringsAsFactors=FALSE,
   Experimental.Method = c("X-Ray", "NMR", "Electron Microscopy", "Other",
                           "Multi Method", "Total"),
              Proteins = c(126880, 11062, 2277, 256, 129, 140604),
         Nucleic.Acids = c(2012, 1279, 31, 4, 5, 3331),
    ProteinComplex = c(6547, 259, 800, 6, 2, 7614),
                 Other = c(8, 8, 0, 13, 1, 30),
                 Total = c(135447, 12608, 3108, 279, 137, 151579)
)



```

>Q2: Type HIV in the PDB website search box on the home page and determine how many
HIV-1 protease structures are in the current PDB?

There are 1157 HIV-1 structures found as of 2019-05-07 See:
http://www.rcsb.org/pdb/results/results.do?tabtoshow=Current&qrid=AC4D3E3E


Do section 2 in VMD 



```{r}
library(bio3d)

pdb <- read.pdb( "1hsg.pdb")
pdb

```
```{r}
#pdb atom dataframe 

pdb$atom

# aa residues from the atom dataframe 
pdb$atom$resid
```

Atom selection is done via the function **atom.select()**

```{r}
?atom.select

prot.pdb <- atom.select(pdb, "protein", value = TRUE)
write.pdb(prot.pdb, file = "1hsg_protein.pdb")
```
 
 So the value xyz is triple , there is 3 coordinates for each atom 

```{r}
#pdb$atom[inds$atom, ]
```

```{r}
lig.pdb <- atom.select(pdb, "ligand", value = TRUE)
write.pdb(lig.pdb, file = "1hsg_ligand.pdb")
```

Adenylate kinase structure analysis from PDB

First get the sequence of Adenylate kinase from 1AKE entry 


```{r}
aa <- get.seq("1ake_A")
```

