---
title: "class_09"
author: "Dominika Kardasz"
date: "April 30th, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Read in the files containing data from biopsies
```{r}
wisc.df <- read.csv("WisconsinCancer.csv")
head(wisc.df)  
```


Convert column 3 to 32 to matrix using **as.matrix** and set rownames to the id numbers 
```{r}
wisc.data <- as.matrix(wisc.df[,3:32])
rownames(wisc.data) <- wisc.df$id

```

Save the diagnosis as a data vector of 1 and 0 values 

```{r}
diagnosis <- as.numeric( wisc.df$diagnosis == "M")

```

>Q1. How many observations are in this dataset?

```{r}
nrow(wisc.df)
```

>Q2. How many variables/features in the data are suffixed with _mean?

```{r}
length( grep( "_mean", colnames(wisc.data)) )

#colnames(wisc.data)
```

>Q3. How many of the observations have a malignant diagnosis?

```{r}
table(wisc.df$diagnosis)
```


Check column means and standard deviations

```{r}

colMeans(wisc.data)

round( apply(wisc.data,2,sd), 1)

```


Perform PCA on your data 

```{r}
wisc.pr <- prcomp( wisc.data, scale= TRUE )

# Look at summary of results
summary(wisc.pr)
```



>Q4. From your results, what proportion of the original variance is captured by the first principal components (PC1)?

```{r}
summary(wisc.pr)
```
>A: 44.27%

>Q5. How many principal components (PCs) are required to describe at least 70% of the original variance in the data?
>A: 3

>Q6. How many principal components (PCs) are required to describe at least 90% of the original variance in the data?
>A: 7

Create a biplot of the wisc.pr using the **biplot()** function
```{r}
biplot(wisc.pr)
```

>Q7. What stands out to you about this plot? Is it easy or difficult to understand? Why?
>A: Hard to understand , trends rather hard to see because Rownames are used as the plotting character for biplots 


```{r}
# Scatter plot observations by components 1 and 2

plot(wisc.pr$x[,1], wisc.pr$x[,2], col=diagnosis+1, xlab = "PC1", ylab = "PC2")
```

>Q8. Generate a similar plot for principal components 1 and 3. What do you notice about these plots?

```{r}
# Repeat for components 1 and 3
plot(wisc.pr$x[,1], wisc.pr$x[,3], col = (diagnosis + 1), 
     xlab = "PC1", ylab = "PC3")
```
>A: Benign data deems to be very similar to each other and the malignant one seems to be more spread. Because principal component 2 explains more variance in the original data than principal component 3, first plot has a cleaner cut separating the two subgroups. Overall, the plots indicate that **principal component 1 is capturing a separation of malignant from benign samples**.

###Variance explained 

Calculate variance of each compomnent:

```{r}
pr.var <- wisc.pr$sdev^2
head(pr.var)
```

Calculate the precentage variance explanined by each principal component:

```{r}
# so you ave data points and divide by the sum of nr of points and then do times 100 for precentage 
pve <- (pr.var/sum(pr.var))*100
head(pve)
```

Plot variance explained for each principal component

```{r}
plot(pve, xlab = "Principal Component", 
     ylab = "Precentage of Variance Explained", 
     ylim = c(0, 100), type = "o")
```

```{r}
# Alternative scree plot of the same data, note data driven y-axis
barplot(pve, ylab = "Precent of Variance Explained",
     names.arg=paste0("PC",1:length(pve)), las=2, axes = FALSE)
axis(2, at=pve, labels=round(pve,2)*100 )
```

Using the **cumsum()** function, create a plot of cumulative proportion of variance explained.


```{r}
# Plot cumulative proportion of variance explained
plot(cumsum(pve), xlab = "Principal Component", 
     ylab = "Cumulative Precentage of Variance Explained", 
     ylim = c(0, 100), type = "o")
```



```{r}
#Setup side by side plots with par(mfcol=c(1,2))
par(mfcol=c(1,2))

plot(pve,  xlab = "Principal Component", 
     ylab = "Precentage of Variance Explained", 
     ylim = c(0, 100), type = "o")


plot(cumsum(pve), xlab = "Principal Component", 
     ylab = "Cumulative Precent of Variance Explained", 
     ylim = c(0, 100), type = "o")
```



```{r}
# Scale the wisc.data data: data.scaled
data.scaled <- scale(wisc.data)
```

```{r}
data.dist <- dist(data.scaled)
```

```{r}
wisc.hclust <- hclust( data.dist, method = "complete")
```

```{r}
plot(wisc.hclust)
abline(h=19, col="red", lty=2)
```

```{r}
wisc.hclust.clusters <- cutree(wisc.hclust, k = 4)
table(wisc.hclust.clusters)
```
we can use the table function to compare the cluster membership to actual diagnoses 

```{r}
table(wisc.hclust.clusters, diagnosis)
```

#Clustering on PCA results 

Using the minimum number of principal components required to describe at least 90% of the variability in the data, create a hierarchical clustering model with the linkage method="ward.D2". We use Ward’s criterion here because it is based on multidimensional variance like principal components analysis. Assign the results to wisc.pr.hclust. 

```{r}
#we have to work with PC1 to PC 7 
wisc.pca.hclust <- hclust(dist(wisc.pr$x[,1:7]) , method="ward.D2")

```


```{r}
plot(wisc.pca.hclust)
```


```{r}
grps <- cutree(wisc.pca.hclust, k=2)
table(grps)
```


```{r}
table(grps, diagnosis)
```


```{r}
#plot and colour with resuts from hierarcal clustering 
plot(wisc.pr$x[,1:2], col=grps)
```


```{r}
plot(wisc.pr$x[,1:2], col=diagnosis+1)

```


```{r}
#call rgl package and make a 3d plot 
library(rgl)
plot3d(wisc.pr$x[,1:3], xlab="PC 1", ylab="PC 2", zlab="PC 3", cex=1.5, size=1, type="s", col=diagnosis+1)

#To include the interactive rgl plot in your HTML renderd lab report you can add the R code

rglwidget(width = 400, height = 400)
```

##7.Prediction

We will use the **predict()** function that will take our PCA model from before and new cancer cell data and project that data onto our PCA space.

```{r}
#url <- "new_samples.csv"
url <- "https://tinyurl.com/new-samples-CSV"
new <- read.csv(url)
npc <- predict(wisc.pr, newdata=new)
npc
```

Plot these new patients on our PCA plot from before  

```{r}
plot(wisc.pr$x[,1:2], col=grps)
points(npc[,1], npc[,2], col="blue", pch=16, cex=3)
text(npc[,1], npc[,2], c(1,2), col = "white")
```
>Q17. Which of these new patients should we prioritize for follow up based on your results?
>A: Patient 2
