---
title: "class 06 r function"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##about 
this is my **class 6** r markdown document with code and notes 

```{r}
plot(1:10, typ="l", col="blue")
```


#read the txt files called test 

```{r}
read.table("test1.txt", header=TRUE, sep=",")
```
```{r}
read.table("test2.txt", header=TRUE, sep="$")
```
```{r}
read.table("test3.txt", header=F, sep="")
```

#our first silly function example 

```{r}
add <- function(x, y=1) {
# Sum the input x and y
x + y
}
```
```{r}
add(x=1, y=4)

```
add(1, 4)
add(1)
#that one will give error 
add( c(1, 2, 3) )
#?
add( c(1, 2, 3), 4 )
add(1, 2, 2)
add(x=1, y=“b”)
## You need a “name”, “arguments” and “body”…
```{r}
rescale <- function(x) {
rng <-range(x)
(x - rng[1]) / (rng[2] - rng[1])
}

```
# Test on a small example where you know the answer
rescale(1:10)


```{r}
rescale(1:10)
```

#na.rm = true change ot and that was about the range r values argumeht 
#should then give anwsers 1 and 10


```{r}

rescale3 <- function(x, na.rm=TRUE, plot=FALSE) {
rng <-range(x, na.rm=na.rm)
print("Hello")
answer <- (x - rng[1]) / (rng[2] - rng[1])
print("is it me you are looking for?")
if(plot) {
plot(answer, typ="b", lwd=4)
}
print("I can see it in ...")
return(answer)
}

```






