---
title: "class7"
output: github_document


---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#source a file froom last day 

```{r}
source("http://tinyurl.com/rescale-R")
```

```{r}
rescale(1:10)
```



#find missing NA values in teo vectors 
```{r}
# Lets define an example x and y
x <- c( 1, 2, NA, 3, NA)
y <- c(NA, 3, NA, 3, 4)
```

```{r}
is.na(x)
```

```{r}
is.na(y)
```
Try putting these together with an AND

```{r}
is.na(x) & is.na(y)
```


Take the sum() to ifnd out how many TRUE values we have and thus how many NAs we had in both x and y 

```{r}
sum( is.na(x)& is.na(y))
```

Now i can make it into our first function ...
```{r}
both_na<- function(x,y){
  sum( is.na(x) & is.na(y) )
}
```

```{r}
both_na(x, c(NA, 3, NA, 2, NA))
```


Test, test , test
```{r}
x <- c(NA, NA, NA)
y1 <- c( 1, NA, NA)
y2 <- c( 1, NA, NA, NA)
```

```{r}
both_na(x, y2)
```

```{r}
y3 <- c(1, NA, NA, NA, NA, NA, NA)
both_na(x,y3)
```

Explanation - the forst number is recycles to match the second vector length , so in the case above the NA gets recyced from x value , the one we compare against i guess 

```{r}
3==3
```
```{r}
3<2
```

```{r}
3!=2
```
```{r}
length(x)
```

```{r}
length(y2)
```

Now lets try the both_na2() function on our different ength input vectors 
```{r}
both_na(x, y2)
```

```{r}
which(c(F,F,T,F,T))


#which(is.na(c(1,2, NA,4)))
```
```{r}
x <- c( 1, 2, NA, 3, NA)
y <- c(NA, 3, NA, 3, 4)
both_na3(x,y)

```
#intersect function
```{r}
df1

```

```{r}
df2

```


#extracting the ID column- make things simple
```{r}
x<-df1$IDs
y<-df2$IDs

x
y
```
```{r}
intersect(x, y)
```

#now i knw that this is the ntersection , but i want it ti output the whole thing with the values from each experimet etc 

```{r}
which(x %in% y)

```

```{r}
x[x %in% y]

```
Use R studio shortcut 'CODE > EXTRACT FUNCTION' to turn our snipett into a working function 

```{r}
#did this by clicling on the code then >code >extract function 
gene_intersect <- function(x, y) {
  cbind( x[x %in% y],
  y[y %in% x])
}
```

```{r}
gene_intersect(df1$IDs, df2$IDs)
```

```{r}
gene_intersect2(df1, df2)
```

```{r}
gene_intersect3(df1,df2)
```
```{r}
merge(df1, df2)
```
find average grade dropping one worst asignment grade
```{r}
#exhercise to determine the average grade dropping lowest score

x <-c(100, 90, 90, 90, 90, 90, 97, 80)

#then find the minimum score 
 min(x)

#then get rid of min score and calc average 
(sum(x)-min(x))/(length(x)-1)

grade<-function(x){
  
  (sum(x)-min(x))/(length(x)-1)
}
 

```




